---
title: "Covid91 vaccine study Final2021"
author: "Fernando D Velasquez Perdomo"
date: "04/30/2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

#Methods  

The First step is we isolate the sub-populations into four different groups as Males, Females, LGBTQ, and Drug users.

```{r}
Maledata<- filter(FinalData,sex=="guy")
Maledata$sex<- as.factor(Maledata$sex)
Maledata$infected=factor(Maledata$infected)
Maledata$infected <- relevel(Maledata$infected, "OK")
```


# Males

## Bar Chart

```{r}

Males <- Maledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
From the Bar Chart we can determined that the treatment has no effect on Males as approximately 57% of males with the treatment and 43% with placebo have a similar infection rate.

## Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=Maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```
  
The table shows that 340 males out of 10,400 were infected as 3.72% who took the treatment were infected as well 2.55% who were given the placebo were as well infected. The vaccine is shown to not be effective for males as 50% who were given the vaccine were likely to be infected in the future.  

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
From what we can gather from the Chi Squared table the p value is 0.0006181 which is highly unlikely that the results occur by chance.  
From the Fisher Exact test we can determined that the vaccine is about 60% effective while the 40% was ineffective.    

$H_0$ The null hypothesis explains that the vaccine does not make a difference on infection rate if a male decides to take the vaccine.  
$H_A$ The alternative hypothesis states that their is a difference in infection rate if a male decides to take the vaccine.     

# Females

```{r}
Femaledata<- filter(FinalData,sex=="gal")
Femaledata$sex<- as.factor(Femaledata$sex)
Femaledata$infected=factor(Femaledata$infected)
Femaledata$infected <- relevel(Femaledata$infected, "OK")
levels(Femaledata$infected)
```

## Bar Chart

```{r}

Females <- Femaledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The Bar Chart shows that approximately out of 607 females infected with covid-91, 32% were given the vaccine and 66% were given the placebo. The Vaccine may be effective on females.



## Numerical Summary
```{r}
table2 <- xtabs(~infected + treatment, data=Femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```
From the table we can see that out of 9,200 females, 607 were infected with covid. Out of those 607 females 4.44% of were given the vaccine and 8.67% were given the placebo. We can conclude that treatment is effective in females.  

## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
The Chi Squared table show us that the p value is so small it scour by chance.  

$H_0$ The null hypothesis explains that the vaccine has no effect in infection rate on females who took the vaccine.  
$H_A$ The alternative hypothesis states that their is a difference in infection rate if a male decides to take the vaccine.  

We reject the null hypothesis as the vaccine is 67% effective on females.  

# LGBTQ

```{r}
LGBTQdata<- filter(FinalData,LGBTQ=="gay")
LGBTQdata$LGBTQ<- as.factor(LGBTQdata$LGBTQ)
LGBTQdata$infected=factor(LGBTQdata$infected)
LGBTQdata$infected <- relevel(LGBTQdata$infected, "OK")
```



## Bar Charts

```{r}

LGBTQ <- LGBTQdata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

From what we can determined from the bar chart out of 116 people who contracted covid about 70% were given the Covid vaccine. We can conclude that the vaccine may not be effective for LGBTQ community. 

## Numerical Summary

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQdata)
rowPerc(table3)
colPerc(table3)
print(table3)
```
  
From the table, we can see that 116 LGBTQ members were infected with covid which 70% were given the vaccine. Compared to the 30% who received the placebo. 

## Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```
From the Chi-squared table we can see that the p value is so small that is unlikely that it occur by chance.  

$H_0$ The null hypothesis explains that the vaccine has no effect in infection rate on LGBTQ who took the vaccine.  
$H_A$ The alternative hypothesis states that their is a difference in infection rate if a LGBTQ decides to take the vaccine.  

The vaccine is 29% effective on the LGBTQ community.  





# Drug Users

```{r}
Druggiedata<- filter(FinalData,DrugUser=="yes")
Druggiedata$DrugUser<- as.factor(Druggiedata$DrugUser)
Druggiedata$infected=factor(Druggiedata$infected)
Druggiedata$infected <- relevel(Druggiedata$infected, "OK")
```


## Bar Charts

```{r}

Druggies <- Druggiedata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
From the Bar Chart approximately we see 117 infected drug users which 32% were given the vaccine.  


## Numerical Summary

```{r}
table4 <- xtabs(~infected + treatment, data=Druggiedata)
rowPerc(table4)
colPerc(table4)
print(table4)
```
  
From the table we can see 117 out of 2,280 were infected with covid which 33.3% were given the vaccine and the other 66% were given the placebo. The drug users who were vaccinated were 60% less likely to be sick from covid.  


## Inferential Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```
From the Chi Squared table we can see that the p value is so small that the results are highly unlikely to occur by chance.  

$H_0$ The null hypothesis explains that the vaccine has no effect in infection rate on Drug users who took the vaccine.  
$H_A$ The alternative hypothesis states that their is a difference in infection rate if a Drug users decides to take the vaccine.  

We can determine that the vaccine has a 60% effectiveness in Drug users.  


# Discussion and Conclusion 

From what we can conclude from looking at 4 different sub-population we can determine that the group with the highest effectiveness treatment/vaccine were the Females with a 67% of effectiveness. Drug users were the second group who reacted well to the vaccine as the vaccine was 60% effective. While the Males and LGBTQ sub population responded poorly to the vaccine as males had a 40% effective rate and LGBTQ community had a 29% effectiveness.  



