---
title: "Covid91 vaccine study Final2021"
author: "Fernando D Velasquez Perdomo"
date: "04/30/2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

#Methods  

The First step is we isolate the sub-populations into four different groups as Males, Females, LGBTQ, and Drug users.

```{r}
Maledata<- filter(FinalData,sex=="guy")
Maledata$sex<- as.factor(Maledata$sex)
Maledata$infected=factor(Maledata$infected)
Maledata$infected <- relevel(Maledata$infected, "OK")
```


# Males

## Bar Chart

```{r}

Males <- Maledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
From the Bar Chart we can determined that the treatment has no effect on Males as approximately 57% of males with the treatment and 43% with placebo have a similar infection rate.

## Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=Maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```
  
The table shows that 340 males out of 10,400 were infected as 3.72% who took the treatment were infected as well 2.55% who were given the placebo were as well infected. The vaccine is shown to not be effective for males as 50% who were given the vaccine were likely to be infected in the future.  

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
From what we can gather from the Chi Squared table the p value is 0.0006181 which is highly unlikely that the results occur by chance.  
From the Fisher Exact test we can determined that the vaccine is about 60% effective while the 40% was ineffective.    

$H_0$ The null hypothesis explains that the vaccine does not make a difference on infection rate if a male decides to take the vaccine.  
$H_A$ The alternative hypothesis states that their is a difference in infection rate if a male decides to take the vaccine.     

# Females

```{r}
Femaledata<- filter(FinalData,sex=="gal")
Femaledata$sex<- as.factor(Femaledata$sex)
Femaledata$infected=factor(Femaledata$infected)
Femaledata$infected <- relevel(Femaledata$infected, "OK")
levels(Femaledata$infected)
```

## Bar Chart

```{r}

Females <- Femaledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The Bar Chart shows that approximately out of 607 females infected with covid-91, 32% were given the vaccine and 66% were given the placebo. The Vaccine may be effective on females.



## Numerical Summary
```{r}
table2 <- xtabs(~infected + treatment, data=Femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```
From the table we can see that out of 9,200 females, 607 were infected with covid. Out of those 607 females 4.44% of were given the vaccine and 8.67% were given the placebo. We can conclude that treatment is effective in females.  

## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
The Chi Squared table show us that the p value is so small it scour by chance.  

$H_0$ The null hypothesis explains that the vaccine has no effect in infection rate on females who took the vaccine.  
$H_A$ The alternative hypothesis states that their is a difference in infection rate if a male decides to take the vaccine.  

We reject the null hypothesis as the vaccine is 67% effective on females.  




# Drug Users






# Drug Users





